---
title: "R_Project"
output: html_document
date: '2022-05-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```


```{r}

data1 <- read.csv("NYPD_Complaint_Data_Historic.csv", na.strings=c(""," ","NA"))
data2 <- read.csv("NYPD_Complaint_Data_Current_Year_To_Date.csv", na.strings=c(""," ","NA"))

# remove unneccessary columns
data1 = data1 %>%
  select(c(1:3,8:14, 16, 24:30, 32:35))

#remove and reorder so it matches above df
data2 = data2 %>%
  select(c(1, 4, 5, 13, 16, 19, 20, 8, 14, 3, 21, 24:27,33:35, 23, 28:30 ))

v1 == v2

dim(data1) #160448
dim(data2) #8843

#combine into one datafram
data = rbind(data1, data2)

dim(data) #169291

data = data[!duplicated(data), ]

```


Ridership Dataset
```{r}
#annual data from: https://new.mta.info/agency/new-york-city-transit/subway-bus-ridership-2020

#total for 2021 

Year = c(2015:2021)
annual_ridership = c(1762565419, 1756814800, 1727366607, 1680060402, 1697787002, 639541029, 761142069)

#below data is from march 2020 - May 2022
data3 <- read.csv("MTA_Daily_Ridership_Data_Beginning_2020.csv", na.strings=c(""," ","NA"))
head(data3$Date)

data3$Date = as.Date(data3$Date, format="%m/%d/%Y")
data3$Year = format(data3$Date, format="%Y")

#get annual data for 2021
data3 %>%
  group_by(Year) %>%
  summarise(sum(Subways..Total.Estimated.Ridership))
data3
ridership = data.frame(Year, annual_ridership)
ridership
```




```{r}
# convert date columns to date types
# create a year, month column from date

data$CMPLNT_FR_DT = as.Date(data$CMPLNT_FR_DT, format="%m/%d/%Y")
class(data$CMPLNT_FR_DT)

data$Year = format(data$CMPLNT_FR_DT, format="%Y")
data$Year_Month = format(data$CMPLNT_FR_DT, format="%Y-%m")
dim(data)
# as.POSIXct(data$CMPLNT_FR_TM, format= "%H:%M")


```


*** revisit ***
```{r}

#how can i fill in empty zip code with lat long data?

sapply(data, function(x) sum(is.na(x)))

data %>%
  filter(is.na(BORO_NM))

# library("ggmap")
# data$Lat_Lon
# revgeocode(data$Lat_Lon, output="more")
```



***Update visuals***

```{r}

annual_incident_prct

ggplot(data=annual_incident_prct, aes(x=Year, y=incident)) + geom_col()


ggplot(data=annual_incident_prct, aes(x=Year, y=incident, fill=incident)) + geom_bar(stat="identity", fill="white", color="red") + labs(x = "Year", y="Count of Incidents") + coord_flip() + expand_limits(y = 12500) + scale_y_continuous(labels = comma) #+ geom_text(aes(label = incident, y = incident), size = 3, position = position_stack(vjust = 1.05))



```

```{r}

annual_incident = data %>%
  group_by(Year) %>%
  summarise(incident = n())

data = data[!(data$Year %in% c('1016','1955', '1966', '1971', '1985', '1990', '2000', '2001', '2003', '2004', '2005', '2006', '2007', '2008', '2009', '2010', '2011', '2012', '2013', '2014', NA)), ]

dim(data)
options(scipen = 999)
ridership$Year = as.character(ridership$Year)
annual_incident_prct = inner_join(annual_incident, ridership, by="Year")
annual_incident_prct
annual_incident_prct = annual_incident_prct %>%
  mutate(incident_prcnt = incident/annual_ridership*100)



annual_incident_prct
(0.0011148620-0.0006991454)/0.0006991454

#2020 Ridership by borough/station data: https://new.mta.info/agency/new-york-city-transit/subway-bus-ridership-2020

```

***update visuals***

This graph shows that there is an increase of subway crime in the past 2 years relative to the amount of people who are using the subway
```{r}
ggplot(data = annual_incident_prct, aes(x=Year, y=incident_prcnt, group=1)) + geom_line(color="red") + ylim(0.0005,0.0015) + labs(x = "\nYear", y="Incident Per Rider\n")
```



Removed 66 rows where borough location was not included
Grouped data by borough to get total incidents per year per borough

```{r}
dim(data) #74513

data = data %>%
  filter(!is.na(BORO_NM)) 
dim(data) #74447

boro_incident = data %>%
  group_by(Year, BORO_NM) %>%
  summarise(incident = n())
boro_incident

```

merged data from total ridership per year per borough to the table above in order to get the incident ratio per the volume of people taking the subway

```{r}

boro_ridership = read.csv("ridership_annual_boro.csv", na.strings=c(""," ","NA"))
boro_ridership$Year = as.character(boro_ridership$Year)
boro_ridership$BORO_NM = toupper(boro_ridership$BORO_NM)

boro_ridership

df_boro = merge(boro_ridership, boro_incident, by=c("Year","BORO_NM"))

df_boro = df_boro %>%
  mutate(incident_prcnt = incident/Ridership*100)

df_boro

```



***come up with a better scale or way to represent these large numbers a better way/abbreviate***

```{r}
library(scales)
class(ridership$annual_ridership)

ggplot(data=ridership, aes(x=Year, y=annual_ridership, fill=annual_ridership)) + geom_bar(stat="identity", fill="white", color="blue") + labs(x = "Year", y="Ridership") + coord_flip() + scale_y_continuous(labels = comma) #+ geom_text(aes(label = annual_ridership, y = annual_ridership), size = 3, position = position_stack(vjust = .5))

ridership

```

This graph shows that there is an increase of subway crime in the past 2 years relative to the amount of people who are using the subway

```{r}
df_boro

ggplot(data = df_boro, aes(x=Year, y=incident_prcnt, group=BORO_NM, color=BORO_NM)) + geom_line() + ylim(0,0.0025) + scale_color_manual(values=c("#097969", "blue", "red", "orange")) + labs(x = "\nYear", y="Incident per Rider\n")

```



```{r}

law_cat_year = data %>%
  group_by(Year, LAW_CAT_CD) %>%
  summarise(count=n())


law_cat_year = law_cat_year %>%
  group_by(Year) %>%
  mutate(Total=sum(count)) %>%
  mutate(Prct = paste0( round(count/Total*100),"%" ) )
  

law_cat_year

```
Distribution of level of crimes per year per total volume

```{r}
ggplot(data=law_cat_year, aes(x=Year, y=count, fill=LAW_CAT_CD)) + geom_bar(stat="identity")

ggplot(data=law_cat_year, aes(x=Year, y=count, fill=LAW_CAT_CD)) + geom_bar(stat="identity") + geom_text(aes(label=Prct), position = position_stack(vjust = 0.5), size = 3) + labs(x = "\nYear", y="Count of Incidents\n", fill="Incident Type")

```



Distribution of level of crimes per year per total volume / For each borough

***add percentages?***
***order facets***

```{r}

boro_law =data %>%
  group_by(Year, BORO_NM, LAW_CAT_CD) %>%
  summarise(count=n())

boro_law = boro_law %>%
  group_by(Year, BORO_NM) %>%
  mutate(Total=sum(count)) %>%
  mutate(Prct = paste0( round(count/Total*100),"%" ) )


boro_law_m = boro_law %>%
  group_by(Year, BORO_NM) %>%
  mutate(Total=sum(count)) %>%
  mutate(Prct = paste0( round(count/Total*100),"%" ) ) %>%
  filter(BORO_NM == "MANHATTAN")

boro_law
boro_law_m

```
***adjust grid width***
Across all boroughs, distribution of misdemeanors increased slightly

```{r}
ggplot(data=boro_law, aes(x=Year, y=count, fill=LAW_CAT_CD)) + geom_bar(stat="identity") + facet_wrap(vars(BORO_NM)) + geom_text(aes(label=Prct), position = position_stack(vjust = 0.5), size = 1.5) #facet_grid(cols=vars(drv))

ggplot(data=boro_law, aes(x=Year, y=count, fill=LAW_CAT_CD)) + geom_bar(stat="identity") + facet_grid(.~BORO_NM) + geom_text(aes(label=Prct), position = position_stack(vjust = 0.5), size = 2) + theme(axis.text.x = element_text(angle=90, hjust=1)) + labs(x = "\nYear", y="Count of Incidents\n", fill="Incident Type") #facet_grid(cols=vars(drv))

# Manhattan only
# ggplot(data=boro_law_m, aes(x=Year, y=count, fill=LAW_CAT_CD)) + geom_bar(stat="identity") + geom_text(aes(label=Prct), position = position_stack(vjust = 0.5), size = 3)

```


```{r}

unique(data$OFNS_DESC)

#update NA offense defense with PD_DESC
data$OFNS_DESC <- ifelse(is.na(data$OFNS_DESC), data$PD_DESC, data$OFNS_DESC)

data %>%
  filter(is.na(LAW_CAT_CD))

unique(data$OFNS_DESC)

data %>%
  group_by(BORO_NM, OFNS_DESC) %>%
  summarise(counts=n()) %>%
  arrange(BORO_NM, desc(counts))


data %>%
  filter(LAW_CAT_CD == "MISDEMEANOR") %>%
  group_by(Year, OFNS_DESC) %>%
  summarise(count=n()) %>%
  arrange(Year, desc(count))

```

***graphically show top 5 crimes somehow other than bar chart***
```{r}
top_crimes = data %>%
  group_by(OFNS_DESC) %>%
  summarise(counts=n()) %>%
  mutate(Prct=paste0( round(counts/74447*100),"%" )) %>%
  arrange(desc(counts)) %>%
  head(5)
  
top_crimes

# top 5 crimes for each borough in past years have been 1)CRIMINAL MISCHIEF & RELATED OF 2)GRAND LARCENY 3)FORGERY 4)ASSAULT 3 & RELATED OFFENSES 5)HARRASSMENT 2
# except for queens Robbery instead of Forgery

```







```{r}

unique(data$OFNS_DESC)


data %>%
  group_by(LAW_CAT_CD, OFNS_DESC) %>%
  summarise(counts = n()) %>%
  arrange(LAW_CAT_CD, desc(counts))
  



```




Mapping for top 5 stations with incidents for each year
***need to complete***
```{r}

# data %>%
#   filter(BORO_NM == "MANHATTAN") %>%
#   group_by(Year, STATION_NAME) %>%
#   summarise(count=n()) %>%
#   top_n(10) %>%
#   arrange(Year,desc(count))

top_manhattan_station = data %>%
  filter(BORO_NM == "MANHATTAN") %>%
  group_by(Year, Lat_Lon) %>%
  summarise(count=n()) %>%
  mutate(Total=sum(count)) %>%
  mutate(Prct=paste0( round(count/Total*100),"%" )) %>%
  arrange(Year, desc(count)) %>%
  do(head(., n = 5))
top_manhattan_station

data_lat_long = data %>%
  distinct(Lat_Lon, .keep_all= TRUE) %>%
  select(Latitude, Longitude, Lat_Lon, TRANSIT_DISTRICT, STATION_NAME)

top_manhattan_station_21 = left_join(top_manhattan_station, data_lat_long,by="Lat_Lon", all.=FALSE)

top_manhattan_station_19 = left_join(top_manhattan_station, data_lat_long,by="Lat_Lon", all.=FALSE)

top_manhattan_station_21 = top_manhattan_station_21 %>%
  filter(Year == 2021)

top_manhattan_station_19 = top_manhattan_station_19 %>%
  filter(Year == 2019)

#test of leaflet mapping with test codes#
library(leaflet)
station = c("station1", "station2", "station3")
lat = c(40.74830)
long = c(-73.98818)
dfr <- as.data.frame(cbind(station,long,lat))
dfr$long = as.numeric(dfr$long)
dfr$lat = as.numeric(dfr$lat)
dfr

leaflet(data=dfr) %>%
  addTiles() %>%
  addMarkers()
  
leaflet(data=dfr) %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 12) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers()
#####################



leaflet(data=top_manhattan_station_21) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers(
    lng = -73.990, lat = 40.757,
    label = "42 ST.-PORT AUTHORITY BUS TERM",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.987, lat = 40.756,
    label = "42 ST.-TIMES SQUARE",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.937, lat = 40.804,
    label = "125 STREET",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.945, lat = 40.808,
    label = "125 STREET",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.991, lat = 40.751,
    label = "34 ST.-PENN STATION",
    labelOptions = labelOptions(noHide = T))

leaflet(data=top_manhattan_station_19) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  # addCircleMarkers() %>%
  addMarkers(
    lng = -73.990, lat = 40.757,
    label = "42 ST.-PORT AUTHORITY BUS TERM",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.937, lat = 40.804,
    label = "125 STREET",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.983, lat = 40.737,
    label = "14 ST.-UNION SQUARE",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.953, lat = 40.811,
    label = "125 STREET",
    labelOptions = labelOptions(noHide = T)) %>%
    addMarkers(
    lng = -73.988	, lat = 40.748,
    label = "34 STREET",
    labelOptions = labelOptions(noHide = T))

top_manhattan_station_21

```



```{r}

dim(data)
data %>%
  filter(BORO_NM == "MANHATTAN") %>%
  group_by(STATION_NAME, Lat_Lon) %>%
  summarise(n())


length(unique(data$STATION_NAME))

data_clean = data %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "1 AVENUE", "(40.73137039, -73.982563257)", Lat_Lon))
data_clean = data %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "1 AVENUE", "(40.73137039, -73.982563257)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "103 STREET" & (Lat_Lon %in% c("(40.789260224, -73.947710001)",	"(40.78926022400003, -73.94771000099996)", 	"(40.796571143000044, -73.96718568699998)", "(40.79032242500005, -73.94768749799994)" )), "(40.790322425, -73.947687498)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "103 STREET" & (Lat_Lon %in% c("(40.796119229, -73.961443369)", "(40.796119229000055, -73.96144336899994)", "(40.796571143, -73.967185687)")), "(40.796119229, -73.961443369)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "103 STREET" & (Lat_Lon %in% c( "(40.79911039400002, -73.96854247999994)", "(40.799118632, -73.968556923)", 	"(40.799118632000045, -73.96855692299994)" )), "(40.799118632, -73.968556923)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "110 ST.-CATHEDRAL PKWY." & (Lat_Lon %in% c( "(40.79607431300008, -73.96706674599994)", "(40.796571143, -73.967185687)", 	"(40.796571143000044, -73.96718568699998)", "(40.80056181400005, -73.95819730799997)" )), "(40.800561814, -73.958197308)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "110 ST.-CATHEDRAL PKWY." & (Lat_Lon %in% c( "(40.80417664400005, -73.96670152799999)", "(40.814850289, -73.956681848)")), "(40.804176644, -73.966701528)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "110 ST.-CENTRAL PARK NORTH", "(40.798841208, -73.952018585)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "110 STREET", "(40.79481676, -73.944408254)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "116 ST.-COLUMBIA UNIVERSITY", "(40.808012927, -73.963903684)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "116 STREET" & (Lat_Lon %in% c( "(40.79481676, -73.944408254)", "(40.798630555000045, -73.94162037399997)", "(40.79863330000006, -73.94162037099994)", "(40.800694331, -73.941109286)", "(40.800694331000045, -73.94110928599997)")), "(40.7986333, -73.941620371)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "116 STREET" & (Lat_Lon %in% c( "(40.80203233000003, -73.94968294499995)")), "(40.80203233, -73.949682945)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "116 STREET" & !(Lat_Lon %in% c("(40.7986333, -73.941620371)", "(40.80203233, -73.949682945)")), "(40.804428073, -73.95537021)", Lat_Lon))

  
  # mutate(Lat_Lon = ifelse(STATION_NAME == "116 STREET" & (Lat_Lon %in% c( "(40.80323023600005, -73.95252472099996)", "	(40.80442807300005, -73.95537020999996)", "(40.808012927, -73.963903684)", "(40.808749904, -73.952585806)", "(40.80874990400008, -73.95258580599993)")), "(40.804428073, -73.95537021)", Lat_Lon))


data_clean = data_clean %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "125 STREET" & (Lat_Lon %in% c("(40.804384046000045, -73.93742166899995)","(40.800694331, -73.941109286)", "(40.800694331000045, -73.94110928599997)")), "(40.804384046, -73.937421669)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "125 STREET" & (Lat_Lon %in% c("(40.807780622000045, -73.94548472699995)")), "(40.807780622, -73.945484727)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "125 STREET" & (Lat_Lon %in% c("(40.81077276700007, -73.95259158999993)")), "(40.810772767, -73.95259159)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "125 STREET" & (Lat_Lon %in% c("(40.814850289, -73.956681848)", "(40.81485028900005, -73.95668184799997)", "(40.815696257, -73.958292572)", "(40.81569625700007, -73.95829257199993)", "(40.81569900300008, -73.95829618299997)", "(40.81573205300003, -73.94542041099999)")), "(40.815696257, -73.958292572)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "135 STREET" & (Lat_Lon %in% c("(40.81411316100008, -73.94085536499993)")), "(40.814113161, -73.940855365)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "135 STREET" & (Lat_Lon %in% c("(40.81712482300003, -73.94800964599993)")), "(40.817124823, -73.948009646)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "137 ST.-CITY COLLEGE", "(40.821392719, -73.954141241)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "14 ST.-UNION SQUARE", "(40.737203099, -73.983272598)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "14 STREET" & (Lat_Lon %in% c("(40.73444264900007, -73.98990537699996)")), "(40.737203099, -73.983272598)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "42 ST.-TIMES SQUARE", "(40.756043535, -73.986949544)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "42 STREET", "(40.754849215, -73.984119991)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "14 STREET" & (Lat_Lon %in% c("(40.737368955, -73.996840424)", "(40.73736895500008, -73.99684042399997)", "(40.738560221, -73.999683896)", "(40.73856022100006, -73.99968389599996)", "\t(40.738560221, -73.999683896)")), "(40.738560221, -73.999683896)", Lat_Lon)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "42 ST.-GRAND CENTRAL", "(40.751449824, -73.976050545)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "34 ST.-HERALD SQ.", "(40.749780113, -73.987780874)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "34 ST.-PENN STATION", "(40.750999048, -73.990624712)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "34 STREET", "(40.749780113, -73.987780874)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "42 ST.-PORT AUTHORITY BUS TERM", "(40.757240531000036, -73.98979363699993)", Lat_Lon))


data_clean %>%
  filter(BORO_NM == "MANHATTAN") %>%
  filter(Year == 2021) %>%
  group_by(STATION_NAME, Lat_Lon) %>%
  summarise(total = n())

data_clean

```



```{r}

# Clean Lat_Lon data only for the top 10 and 10 lowest stations


data_m_top = data %>% 
  mutate(STATION_NAME = recode(STATION_NAME, "42 ST.-PORT AUTHORITY BUS TERM" = "42 ST.-TIMES SQUARE/PORT AUTHORITY", "42 ST.-TIMES SQUARE" = "42 ST.-TIMES SQUARE/PORT AUTHORITY")) %>% #times square/port authority
  mutate(Lat_Lon = ifelse(STATION_NAME == "42 ST.-TIMES SQUARE/PORT AUTHORITY", "(40.75604353500006, -73.98694954399997)", Lat_Lon)) %>% #times square/port authority
  mutate(Lat_Lon = ifelse(STATION_NAME == "42 ST.-GRAND CENTRAL", "(40.751449824000076, -73.97605054499998)", Lat_Lon)) %>% #group grand central coordinates
  mutate(Lat_Lon = ifelse(STATION_NAME == "34 ST.-HERALD SQ.", "(40.749796581000055, -73.98777004399994)", Lat_Lon)) %>% #group Herald Square Coordinates
  mutate(STATION_NAME = recode(STATION_NAME, "UNION SQUARE" = "14 ST.-UNION SQUARE")) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "14 ST.-UNION SQUARE", "(40.73444264900007, -73.98990537699996)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "FULTON STREET", "(40.71033547800005, -74.00954559999997)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "34 ST.-PENN STATION", "(40.75099904800004, -73.99062471199994)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "59 ST.-COLUMBUS CIRCLE", "(40.76502407900006, -73.98483594999993)", Lat_Lon)) %>%
  mutate(STATION_NAME = recode(STATION_NAME, "PARK PLACE" = "CHAMBERS ST.-WORLD TRADE CENTE")) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "CHAMBERS ST.-WORLD TRADE CENTE", "(40.71154309400004, -74.01046557799998)", Lat_Lon)) %>%
  mutate(STATION_NAME = recode(STATION_NAME, "LEXINGTON AVE." = "LEXINGTON AVE/59 ST", "59 STREET" = "LEXINGTON AVE/59 ST")) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "LEXINGTON AVE/59 ST", "(40.762234771000074, -73.96818808299997)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "47-50 STS./ROCKEFELLER CTR.", "(40.75803003100003, -73.98179465299995)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "215 STREET", "(40.869559485000025, -73.91518555599998)", Lat_Lon)) %>% #lowest volume trains below
  mutate(Lat_Lon = ifelse(STATION_NAME == "155 STREET" & !(Lat_Lon %in% c("(40.83090011900003, -73.94136802099996)")), "(40.82917153600005, -73.93726827299997)", Lat_Lon)) %>%
  
  mutate(STATION_NAME = ifelse(STATION_NAME == "155 STREET" & (Lat_Lon %in% c("(40.82917153600005, -73.93726827299997)")), "155 STREET B/D", STATION_NAME)) %>%
  mutate(STATION_NAME = ifelse(STATION_NAME == "155 STREET" & (Lat_Lon %in% c("(40.83090011900003, -73.94136802099996)")), "155 STREET C", STATION_NAME)) %>%
  
  mutate(Lat_Lon = ifelse(STATION_NAME == "145 STREET" & (Lat_Lon %in% c("(40.82044551200005, -73.93623234699999)", "(40.82044551000007, -73.93622873399994)")), "(40.82044551000007, -73.93622873399994)", Lat_Lon)) %>%
  mutate(STATION_NAME = ifelse(STATION_NAME == "145 STREET" & (Lat_Lon %in% c("(40.82044551000007, -73.93622873399994)")), "145 STREET 3", STATION_NAME)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "BOWERY", "(40.72033214100002, -73.994038162)", Lat_Lon)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "CITY HALL", "(40.71306394100003, -74.00722658699993)", Lat_Lon)) %>%
  mutate(STATION_NAME = ifelse(STATION_NAME == "CANAL STREET" & (Lat_Lon %in% c("(40.722558122000066, -74.00630044499997)", "(40.72259380100007, -74.00635456399993)")), "CANAL STREET 1", STATION_NAME)) %>%
  mutate(Lat_Lon = ifelse(STATION_NAME == "CANAL STREET 1" & (Lat_Lon %in% c("(40.722558122000066, -74.00630044499997)", "(40.72259380100007, -74.00635456399993)")), "(40.72259380100007, -74.00635456399993)", Lat_Lon))

  
data_m_top %>%
  filter(BORO_NM == "MANHATTAN") %>%
  filter(Year == 2020) %>%
  group_by(STATION_NAME, Lat_Lon) %>%
  summarise(total = n())

data %>%
  filter(BORO_NM == "MANHATTAN") %>%
  filter(Year == 2020) %>%
  group_by(STATION_NAME, Lat_Lon) %>%
  summarise(total = n())

data_m_top = data_m_top %>%
  filter(BORO_NM == "MANHATTAN") %>%
  filter(Year == 2020) %>%
  filter(STATION_NAME %in% c("42 ST.-TIMES SQUARE/PORT AUTHORITY", "42 ST.-GRAND CENTRAL", "34 ST.-HERALD SQ.", "14 ST.-UNION SQUARE", "FULTON STREET", "34 ST.-PENN STATION", "59 ST.-COLUMBUS CIRCLE", "LEXINGTON AVE/59 ST", "PARK PLACE", "CHAMBERS ST.-WORLD TRADE CENTE", "215 STREET", "155 STREET C", "155 STREET B/D", "145 STREET 3", "BOWERY", "CITY HALL", "CANAL STREET 1", "191 STREET", "148 ST.-HARLEM", "FRANKLIN STREET", "47-50 STS./ROCKEFELLER CTR.")) %>%
  group_by(STATION_NAME, Lat_Lon) %>%
  summarise(total = n()) %>%
  arrange(desc(total))


#Ridership data for the 20 stations above source https://new.mta.info/coronavirus/ridership

ridership_m_top1 = c("42 ST.-TIMES SQUARE/PORT AUTHORITY", "42 ST.-GRAND CENTRAL", "34 ST.-HERALD SQ.", "14 ST.-UNION SQUARE", "FULTON STREET", "34 ST.-PENN STATION", "59 ST.-COLUMBUS CIRCLE", "LEXINGTON AVE/59 ST", "CHAMBERS ST.-WORLD TRADE CENTE", "47-50 STS./ROCKEFELLER CTR.", "215 STREET", "155 STREET C", "145 STREET 3", "191 STREET", "BOWERY", "148 ST.-HARLEM", "FRANKLIN STREET", "CITY HALL", "155 STREET B/D", "CANAL STREET 1")
ridership_m_top2 = c(20341240, 13162826, 12826841, 10830712, 8855302, 16114281, 7618925, 6273480, 5930847, 5826100, 212402, 367255, 438388, 453577, 454326, 473937, 501136, 536758, 546625, 560302)

length(ridership_m_top1)
length(ridership_m_top2)

data_m_highlow = data.frame(STATION_NAME = ridership_m_top1, RIDERSHIP = ridership_m_top2)


data_m_highlow = merge(data_m_top, data_m_highlow, by="STATION_NAME")
data_m_highlow
data_m_highlow = data_m_highlow %>%
  mutate(Prct=paste0( round(total/RIDERSHIP*100, digit=5),"%" ))

data_m_highlow %>%
  arrange(desc(Prct))
data_m_highlow

#%>%
  #ggplot(aes(x=RIDERSHIP, y=Prct)) + geom_point()
options(ggrepel.max.overlaps = Inf)
install.packages("ggrepel")
library(ggrepel)

ggplot(data = data_m_highlow, aes(x=RIDERSHIP, y=total)) + geom_point(aes(color = "red")) + scale_x_continuous(labels = comma) + labs(x = "\nAnnual Ridership Volume", y="Count of Incidents\n") + theme(legend.position="none")  + theme(legend.position="none")

ggplot(data = data_m_highlow, aes(x=RIDERSHIP, y=total)) + geom_point(aes(color = "red")) + scale_x_continuous(labels = comma) + labs(x = "\nAnnual Ridership Volume", y="Count of Incidents\n") + theme(legend.position="none")  + theme(legend.position="none") + geom_label_repel(aes(label = STATION_NAME),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  size= 1.8,
                  segment.color = 'grey50')

ggplot(data = data_m_highlow, aes(x=RIDERSHIP, y=Prct)) + geom_point(aes(color = "red")) + scale_x_continuous(labels = comma) + labs(x = "\nAnnual Ridership Volume", y="% Incident per Ridership\n") + theme(legend.position="none") + geom_label_repel(aes(label = STATION_NAME),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  size= 1.8,
                  segment.color = 'grey50')

```



```{r}
# install.packages("hms")

substring(data$CMPLNT_FR_TM, 0,2)

incidenttime_2021 = data %>%
  mutate(Hour = substring(data$CMPLNT_FR_TM, 0,2)) %>%
  filter(Year == "2021") %>%
  group_by(Hour) %>%
  summarise(Total = n())


incidenttime_all = data %>%
  mutate(Hour = substring(data$CMPLNT_FR_TM, 0,2)) %>%
  group_by(Year, Hour) %>%
  summarise(Total = n())

ggplot(data = incidenttime_all, aes(x=Hour, y=Total, group=Year, color=Year)) + geom_line() + labs(x = "\nHour of Day", y="% Count of Incidents\n")

incidenttime_2021

ggplot(data = incidenttime_2021, aes(x=Hour, y=Total, group=1)) + geom_line()
ggplot(data = annual_incident_prct, aes(x=Year, y=incident_prcnt, group=1)) + geom_line() + ylim(0.0005,0.0015)
```


```{r}

#impact of ridership before, during, after week of 4/12 shooting on subway
ridership_incident1 = data3 %>%
  select(1:2) %>%
  filter(Date >= "2022-04-5" & Date <= "2022-04-25") %>%
  mutate(Week = case_when(
    Date >= "2022-04-5" & Date <= "2022-04-11" ~ "Week Before",
    Date >= "2022-04-12" & Date <= "2022-04-18" ~ "Week Of",
    Date >= "2022-04-19" & Date <= "2022-04-25" ~ "Week After"
    ))

ridership_incident1
ridership_incident1$Week = factor(ridership_incident1$Week, c("Week Before", "Week Of", "Week After"))
ggplot(data = ridership_incident1, aes(x = Week, Subways..Total.Estimated.Ridership, y = Subways..Total.Estimated.Ridership)) + geom_boxplot() + labs(x = "\nApril 12th Incident", y="Ridership Volume\n")
```

```{r}

#impact of ridership before, during, after week of 1/15 incident
ridership_incident2 = data3 %>%
  select(1:2) %>%
  filter(Date >= "2022-01-8" & Date <= "2022-01-28") %>%
  mutate(Week = case_when(
    Date >= "2022-01-8" & Date <= "2022-01-14" ~ "Week Before",
    Date >= "2022-01-15" & Date <= "2022-01-21" ~ "Week Of",
    Date >= "2022-01-22" & Date <= "2022-01-28" ~ "Week After"
    ))

ridership_incident2$Week = factor(ridership_incident2$Week, c("Week Before", "Week Of", "Week After"))
ggplot(data = ridership_incident2, aes(x = Week, Subways..Total.Estimated.Ridership, y = Subways..Total.Estimated.Ridership)) + geom_boxplot() + labs(x = "\nJanuary 15th Incident", y="Ridership Volume\n")

ridership_incident2
```

